Scaffolder - Software for Microbial Genome Scaffolding.

### 2011-02-17 src/article.tex
Revision of manuscript for readability

  \paragraph*{Results:} This software allows a genome build to be defined in
  terms of the constituent contigs using a relatively simple to write and edit
  scaffold file syntax. This syntax further allows  unknown regions to be
  defined, and inserts to be added to the scaffold to fill gaps. Defining the
  genome construction in a file makes the scaffolding process reproducible and
  comparatively easier to edit and compare than fasta sequence.

@@

High throughput sequencing can produce hundreds of thousands to millions of
sequence reads from a genome. Each sequence represents the machine-read
nucleotides from a short region in the genome. At the time of writing, high
throughput sequencing is limited to producing reads less than $>$1000
nucleotides in length. Therefore to determine a sequence longer than this, such
as a complete genome, these numerous individual reads must be pieced together.
The process of joining reads into longer sequences is the `assembly' stage in
a genome project \cite{miller2010}. \pb

Assembly software takes the nucleotide reads produced by the sequencing
hardware and, in the ideal case, outputs a complete genome sequence compose of
these individual reads. An analogy for this process is a jigsaw puzzle. Each
nucleotide read represents a single piece and the final genome sequence is the
completed puzzle. Sequences of repetitive nucleotide `repeat' regions within
the genome, or biased and incomplete sequencing data may however prevent the
genome being assembled into a continuous sequence. This is analogous to missing
jigsaw pieces or pieces that fit to multiple other pieces. \pb

The recent advent of these current high throughput sequencing methods has lead
to a renewed interest into algorithms for solving the problem of genome
assembly \cite{pop2008,pop2009}. Genome assembly software may however be unable
to produce a complete sequence from these reads. In many cases assembly
algorithms will generate several large contiguous regions of sequence
(`contigs') composed of many individual reads given the available data and
assembly algorithm. These contigs represent the assembly of reads into the
longest possible sequences. These contigs represent a fragmented picture of the
genome and require additional work to join together into larger sequence
lengths. \pb

The process of ameliorating a draft sequence into a finished continuous
sequence can be expensive in both time and laboratory effort. Therefore the
genomic data present in a set of contigs may often be enough for many research
questions \cite{branscomb2002}. A continuous high-quality `finished' genome
sequence does however provide a greater depth of information, such as complete
resolution of repeat regions and more precise estimates of distances between
genomic elements \cite{parkhill2002,fraser2002}. This process of joining
contigs together to form a continuous sequence may be called the `scaffolding'
or `finishing' stage. \pb

@@

Scaffolding is the process of joining a series disconnected contigs into
complete continuous sequence. Due to genomic complexity and missing data,
scaffolding may not ultimately produce a final completed sequence, but may
still succeed in joining a subset of contigs together or filling gaps. An
overview of the scaffolding process is outlined in the following section.

@@

The sequencing process generates reads from either strand of the DNA helix.
Therefore the resulting contigs constructed from these reads may point in
either direction of the DNA strand .The orientation of contigs to point in the
same direction and represent the same DNA strand requires reverse complementing
sequences where necessary. In the case of microbes this orientation may be the
5'-3' strand leading away from the origin of replication.

@@

Contig ordering determines the placement of observed contigs to best represent
their order in the true genome sequence. Correctly placing each contig also
highlights any extra-genomic DNA, such as plasmids, and should be scaffolded
separately from genomic sequence. The contig order is commonly referenced
against the origin of replication where the containing contig is split at the
replication start point. All subsequent contigs are then ordered in the
direction of DNA replication.

@@

Given the correct order and orientation, determining the distances between
contigs results in an estimate of the complete genome size. The size of inter
contig gaps represents the length of an unknown region in the genome. Filling
these regions with unknown nucleotide characters (`N') allows a draft
continuous representation of the genome. This draft scaffold sequence is useful
for representing both the sequenced and unknown regions in the genome.

@@

During the scaffolding process closing and filling gaps between contigs
improves the accuracy of the genome scaffold. This may require returning to the
laboratory to perform additional sequencing or using computational methods of
estimating the gap sequence. This additional sequence is used to replace the
gap between two contigs joining them into a single sequence if they are not
already scaffolded together. Once all contigs have been joined and gaps in
a scaffold closed the genome may be considered finished.

@@

The process of resolving a genome sequence and building a scaffold uses
\emph{in vitro} methods, \emph{in silico} methods, or a combination of both. An
example of a computational method might use paired-read data, where the
sequencing reads were generated at a known distance apart in the genome. The
occurrence of a pair of reds in separate contigs this can be used to estimate
the order and distance. Laboratory methods may use PCR to amplify the unknown
DNA in a gap region, then use traditional Sanger sequencing to determine the
sequence. Computation methods are more preferable as they are less costly in
laboratory time and materials compared to manual gap resolution
\cite{nagarajan2010}. This need for computational scaffolding methods has
resulted in a variety of algorithmic approaches. \pb

Synteny methods compare the assembled contigs to a more complete reference
genome sequence searching for ares of sequence similarity. Areas of
corresponding sequence between the the two sets are can be used to infer contig
organisation \cite{richter2007,zhao2008}. Recombination between the two genomes
can however reduce the effectiveness of this analysis. \pb

Repeat regions may be responsible for many of the gaps in building a genome
sequence. The existence of tandemly repeated nucleotide regions produces many
reads of similar sequence. The assembly of reads generated from repeats can
result in the assemble collapsing into a single repeat instance or ignored by
conservative assembly algorithms. These regions may be resolved by using
reassembly algorithms specifically for repeat resolution
\cite{mulyukov2002,koren2010}. Similarly a related approach may use uniquely
matching reads to try and bridge across gaps between contigs \cite{tsai2010}.
\pb

As mentioned above, the genome may be prepared prior to sequencing so that
reads are produced in pairs at a known distance. This paired-read data
provides an extra level of information on how contigs may be scaffolded
together. Heuristic graph algorithms may take advantage of this data to search
for optimal configuration of contigs in the scaffold using paired-read distances
\cite{dayarian2010}. Additional data using a reference genome may also be
combined when estimating the best contig configuration \cite{pop2004}. \pb

Finally when the scaffold cannot be completely resolved \emph{in silico}
software packages exist to suggest the likely primers necessary for determining
the sequence in gap regions \cite{gordon2001,nagarajan2010}. \pb

@@

Taken together these methods provide a wide variety of possible approaches for
scaffolding a genome. The scaffold may still however require manual editing and
addition of sequence if computational methods are not completely successful.
This involves moving and joining larges blocks of nucleotide text together by
hand which may be expected to be error-prone and precludes any reproducibility.
\pb

The software outlined here ``Scaffolder'' is aimed at creating a syntax and
software framework for producing and editing a genome scaffold. Scaffolder uses
a specific file format to define how contigs are joined, additional sequences
inserted, and for the specification of unknown regions. This syntax allows
a scaffold to be edited in a simpler way and allows the scaffolding process to
be both reproducible and deterministic. \pb

@@

The Scaffolder source code is documented using the Yard library \cite{yard}.
Unix manual pages for the command line were generated using the Ronn library
\cite{ronn}. The manipulation of biological sequences in Scaffolder is
performed using the BioRuby library \cite{goto2010}. A full list of the
software dependencies in Scaffolder can be found in the Gemfile in the root of
each source code directory. \pb

@@

As described in the Background section, the initial sequencing data assembly
may result in an incomplete genome build. The use of computational software and
generation of additional sequence using PCR translates into genome scaffolding
as an on-going process. Therefore the scaffold file should be simple to update
manually in addition to being computationally tractable. This requirement was
therefore the second reason YAML was selected, where YAML is human-readable and
easy to edit in a standard text editor. \pb

The scaffold file takes the form of a list of entries. Each entry corresponds
to a region of sequence, most likely a contig or previously scaffolded set of
contigs. The Scaffolder software processes this list in order, converting each
to the corresponding nucleotide sequence. These nucleotide sequences are then
concatenated to produce a draft super sequence. Each entry in the scaffold file
may have attributes that define whether a sub-sequence or the reverse
complement of the sequence should be used. The types of attributes available
and an example scaffold file are outlined in the Results and Discussion section
below.  \pb

@@

The Scaffolder software outlined here allows computational joining of
nucleotide sequences together into a single contiguous super sequence.
Plain-text scaffold files written in the YAML syntax specify how these
sequences should be joined. This scaffold file is read by the scaffolder
software and used to generate the scaffold sequence. In addition to specifying
which contigs, the scaffold file allows the contigs to be edited to produce
smaller sub-sequences or reverse complemented if necessary. \pb

The process of genome finishing may involve producing additional
oligonucleotide sequence to fill areas of unknown regions in the scaffold. The
scaffold file format allows these additional insert sequences to be added to
bridge these gaps. These inserts can also be treated in a similar way to larger
contig sequences: trimmed and/or reverse complemented to match the
corresponding gap region. \pb

@@

The scaffold file is written using the YAML syntax. YAML is a data format
designed to be both machine readable and easy to edit manually. An example
scaffold file is shown in Figure 1. This file illustrates the text attributes
used to describe a scaffold and the joining of sequences in the genome build.
The basic layout of the scaffold file is a list - a series of entries where
each entry corresponds to a region of sequence in the generated scaffold super
sequence. The right hand side indicates hows the entries are joined together.
This figure is for illustrative purposes and not to scale. \pb

@@

The first line of the scaffold file begins with three dashes. This is
a requirement to indicate the start of a YAML formatted document. The first
entry in the scaffold begins on the second line. The first line of this entry,
highlighted in green, begins with a dash character `-', dashes are used to
denote entries in a YAML list. The first line of each entry in the scaffold
therefore begins with a dash line. The next line is indented by two spaces
- whitespace is required to group similar attributes together. Lines indented
by whitespace are part to the first non-indented preceding line. \pb

@@

The second entry in the scaffold, highlighted in orange, is identified by the
``unresolved'' tag, indicating a region of unknown sequence but known length.
The line below specifies the size of this unknown region. This entry therefore
inserts a region of 20 `N' characters in the scaffold. \pb

@@

This entry includes the ``inserts'' tag to update this sequence with regions of
sequence. These inserts are also added as a YAML list with each insert
beginning with a dash line. The first insert, shown in purple, illustrates
using similar attributes to that of a sequence entry. The reverse, start and
stop tags, are used to trim and reverse complement the insert. Similarly the
`source' tag identifies the corresponding fasta sequence. The ``open'' and
``close'' tags are specific to inserts and determine where the insert is added
in the enclosing sequence. The region of the sequence inside these coordinates
inclusively is replaced with the specified insert fasta sequence. This is
highlighted in the putative scaffold in Figure 1 by the black lines bisecting
the blue sequence. \pb

The next insert, shown in brown, is specified using only the open position tag.
This illustrates that only one of either `open' or `close' tags is required.
When only one of these tag is used the corresponding opposing coordinate
position is calculated from the length in the insert fasta sequence. This
allows inserts to bridge into, and partially fill gap regions with requiring an
end co-ordinate position. \pb

@@

\paragraph{Reverse sort inserts:} If the sequence entry contains a list of
inserts, these inserts are reverse sorted by the insert close coordinate. The
insert with the largest close position is first in the list and the insert
with the smallest close position is last. This preserves the original intended
coordinates for each insert during the process of adding inserts to the
sequence. \pb

@@

Scaffolder relies on both the Ruby programming language and the RubyGems
package management system. These are either already present or easily installed
on recent distributions of Mac OS X or Linux. The presence of both these tools
can be tested by typing the following on the command line. \pb

@@

This code snippet illustrates that the scaffold object is initialised with two
arguments: the YAML loaded scaffold and the location of the fasta sequence
file. The scaffold is an enumerable array and the penultimate line illustrates
the trivial code required to generate the scaffold sequence. Further details of
the classes and methods related to each scaffold entry can be found in the
documentation for the Scaffolder::Region class on the Scaffolder website. \pb

@@

The aims of Scaffolder are twofold. The first is providing software which is
easy to install and simplifies the task of genome finishing. The second aim is
to facilitate reproducibility when joining large nucleotide sequences together
to create a genome scaffold. \pb

@@

Scaffolder is designed to be as simple to use as possible. Assuming the Ruby
and RubyGems software are present, Scaffolder can be installed in a single
command line step. This should negate any possible barriers from requiring
manual compilation of the source code. \pb

Scaffolder uses a minimal and compact syntax to described the genome scaffold.
This syntax is simple to construct and edit whilst being succinct and readable.
The goal of this syntax is to make genome scaffolds easy to construct in
a common text editor. Furthermore Scaffolder provides extensive documentation
both for the CLI and for the Ruby API. Therefore should any part of the
scaffolding process be unclear the documentation should make an solution easy
to find. \pb

@@

Configuring the final sequence in the scaffold file means the build is easier
to edit once already constructed. The YAML text format also allows comparison
of differences between scaffold builds using standard Unix tools such as diff.
This makes scaffold files easier to store in version control systems and allows
genome finishers methods similar to that of software development. \pb

@@

Scaffolder follows the Unix philosophy of ``Do one thing and do it well''.
Scaffolder focuses on creating a simple to use and install tool for creating
a genome scaffold. The YAML data format for creating the scaffold file is
standardised and easily manipulated programmatically. This thereby means the
scaffolding process follows a second Unix tenet: ``If your data structures are
good enough, the algorithm to manipulate them should be trivial.''

@@

  \begin{description}
    \item[API:] Application programming interface
    \item[CLI:] Command line interface
    \item[YAML:] YAML Ain't Markup Language\cite{yaml}
  \end{description}

@@

An example scaffold file written using the YAML syntax \cite{yaml} (left) and
the resulting genome build (right). The scaffold contains three entries and two
inserts. Each entry in the scaffold file text is delimited by a `-' on a new
line and highlighted using separate colours. The scaffold diagram on the right
illustrates how the scaffold sequences are joined and is not to scale. \pb

>>>>

  \paragraph*{Results:} This software allows a genome build to be defined in
  terms of the constituent contigs using a relatively simple to write and edit
  scaffold file syntax. This syntax further allows  unknown regions to be
  defined, and inserts to be added to the scaffold to fill gaps. Defining the
  genome construction in a file makes the scaffolding process reproducible and
  easier to edit and compare than fasta sequence.

@@

High throughput sequencing can produce hundreds of thousands to millions of
sequence reads from a genome. Each sequence represents the machine-read
nucleotides from a short region in the genome. At the time of writing, high
throughput sequencing is limited to producing reads less than 1000 nucleotides
in length. Therefore to determine a sequence longer than this, such as
a complete genome, these numerous individual reads must be pieced together. The
process of joining reads into longer sequences is the `assembly' stage in
a genome project \cite{miller2010}. \pb

Assembly software takes the nucleotide reads produced by the sequencing
hardware and, in the ideal case, outputs a complete genome sequence composed of
these individual reads. An analogy for this process is a jigsaw puzzle. Each
nucleotide read represents a single piece and the final genome sequence is the
completed puzzle. Sequences of repetitive nucleotide `repeat' regions within
the genome, or biased and incomplete sequencing data may however prevent the
genome being assembled into a continuous sequence. This is analogous to missing
jigsaw pieces or pieces that fit to multiple other pieces. \pb

The recent advent of these current high throughput sequencing methods has lead
to a renewed interest into algorithms for solving the problem of genome
assembly \cite{pop2008,pop2009}. Genome assembly software may however be unable
to produce a complete sequence from these reads. In many cases assembly
algorithms will generate several large contiguous regions of sequence
(`contigs') composed of many individual reads. These contigs represent the
assembly of reads into the longest possible sequences given the available data
and assembly algorithm. These contigs represent a fragmented picture of the
genome and require additional work to join together into larger sequence
lengths. \pb

The process of ameliorating a draft sequence into a finished continuous
sequence can be expensive in both time and laboratory effort. Therefore the
genomic data present in a set of contigs may often be enough for many research
questions \cite{branscomb2002}. A continuous high-quality `finished' genome
sequence does however provide a greater depth of information, such as complete
resolution of repeat regions and precise estimates of distances between genomic
elements \cite{parkhill2002,fraser2002}. This process of joining contigs
together to form a continuous sequence may be called the `scaffolding' or
`finishing' stage. \pb

@@

Scaffolding is the process of joining a series disconnected contigs into
complete continuous sequence. Due to genomic complexity and missing data,
scaffolding may not ultimately produce a final completed sequence, but may
still succeed in joining a subset of contigs together or filling gaps. An
overview of the scaffolding process is outlined below.

@@

The sequencing process generates reads from either strand of the DNA helix.
Therefore the resulting contigs constructed from these reads may point in
either direction of the DNA. The orientation of contigs to point in the same
direction and represent the same strand requires reverse complementing
sequences where necessary. In the case of microbes this orientation may be to
the 5'-3' direction leading away from the origin of replication.

@@

Contig ordering determines the placement of observed contigs to best represent
their order in the true genome sequence. Correctly placing each contig also
highlights any extra-genomic DNA, such as plasmids which should be scaffolded
separately from the genomic sequence. Contig order is commonly started at the
origin of replication where the first contig is split at the replication start
point. All subsequent contigs are then ordered in the direction of DNA
replication.

@@

Given the correct order and orientation, determining the distances between
contigs results in an estimate of the complete genome size. The size of inter
contig gaps represents the length of an unknown region in the genome. Filling
these regions with unknown nucleotide characters (`N') allows a draft
continuous sequence. This draft scaffold sequence is useful for representing
both the sequenced and unknown regions in the genome.

@@

During the scaffolding process closing and filling gaps between contigs
improves the completeness of the genome scaffold. Closing gaps may require
returning to the laboratory to perform additional sequencing or using
computational methods of estimating the unknown sequence. This additional
sequence is used to replace the gap between two contigs joining them into
a single sequence if they are not already scaffolded together. Once all contigs
have been joined and gaps in a scaffold closed the genome may be considered
finished.

@@

The process of resolving a genome sequence and building a scaffold uses
\emph{in vitro} methods, \emph{in silico} methods, or a combination of both. An
example of a computational method might use paired-read data, where the
sequencing reads are produced in pairs at a known distance apart in the genome.
The occurrence of a pair of reads in separate contigs can be used to estimate
the order and distance between the contigs. Alternatively laboratory methods
may use PCR to amplify the unknown DNA in a gap region, then use traditional
Sanger sequencing to determine the sequence. Computational methods are more
preferable as they are less costly in laboratory time and materials compared to
manual gap resolution \cite{nagarajan2010}. This need for computational
scaffolding methods has resulted in a variety of algorithmic approaches. \pb

Synteny methods compare the assembled contigs to a more complete reference
genome sequence searching for areas of sequence similarity between the two.
Areas of corresponding sequence between the the two sets are can be used to
infer contig placement and build the scaffold \cite{richter2007,zhao2008}.
Recombination between the two genomes can however reduce the effectiveness of
this analysis. \pb

Repeat regions may be responsible for many of the gaps in building a genome
sequence. The existence of tandemly repeated nucleotide regions produces many
similar reads in the sequence process. The assembly of reads generated from
repeat regions can result in the assemble collapsing into a single repeat
instance or ignored by conservative assembly algorithms. These regions may be
resolved by using reassembly algorithms specifically for repeat resolution
\cite{mulyukov2002,koren2010}. Similarly a related approach may use uniquely
matching reads to try and bridge across gaps between contigs \cite{tsai2010}.
\pb

As mentioned above, the genome may be prepared prior to sequencing so that
reads are produced in pairs at a known distance. This paired-read data
provides an extra level of information on how contigs may be scaffolded
together. Heuristic graph algorithms may take advantage of this data to search
for optimal configuration of contigs in the scaffold using paired-read distances
\cite{dayarian2010}. Additional data using a reference genome may also be
combined when estimating the best contig configuration \cite{pop2004}. Finally
when the scaffold cannot be completely resolved \emph{in silico} software
packages exist to suggest the likely primers necessary for PCR amplifying the
sequence in gap regions \cite{gordon2001,nagarajan2010}. \pb

@@

Taken together these methods provide a wide variety of possible approaches for
scaffolding a genome. The scaffold may still however require manual editing and
addition of sequence if computational methods are not completely successful.
This involves moving and joining larges blocks of nucleotide text together by
hand which may be error-prone and precludes any reproducibility.
\pb

The software outlined here ``Scaffolder'' is aimed at creating a syntax and
software framework for producing and editing a genome scaffold. Scaffolder uses
a specific file format to define how contigs are joined, additional sequences
inserted, and for the specification of unknown regions. This syntax allows
a scaffold to be updated by simply editing the scaffold file and allows the
scaffolding process to be both reproducible and deterministic. \pb

@@

The Scaffolder source code is documented using the Yard library \cite{yard}.
Unix manual pages for the command line were generated using the Ronn library
\cite{ronn}. The manipulation of biological sequences in Scaffolder uses the
BioRuby library \cite{goto2010}. A full list of the software dependencies in
Scaffolder can be found in the Gemfile in the root of each source code
directory. \pb

@@

As described in the Background section, the initial sequencing data assembly
may result in an incomplete genome build. The use of computational software and
generation of additional sequence using PCR means genome scaffolding may be an
on-going process. Therefore the scaffold file should be simple to update
manually in addition to being computationally tractable. This requirement was
therefore the second reason YAML was selected, where YAML is human-readable and
simple to edit in a standard text editor. \pb

The scaffold file takes the form of a list of entries. Each entry corresponds
to a region of sequence, most likely a contig or previously scaffolded set of
contigs. The Scaffolder software processes this list in order, converting each
to the corresponding nucleotide sequence. These nucleotide sequences are then
concatenated to produce a draft super sequence. Each entry in the scaffold file
may have attributes that define whether a sub-sequence or the reverse
complement of the sequence should be used. The types of attributes available,
and an example scaffold file are outlined in the Results and Discussion section
below.  \pb

@@

The Scaffolder software outlined here allows computational joining of
nucleotide sequences together into a single contiguous super sequence.
Plain-text scaffold files written in the YAML syntax specify how these
sequences should be joined. This scaffold file is read by the scaffolder
software and used to generate the scaffold sequence. In addition to specifying
which contigs are required, the scaffold file allows the contigs to be edited
to produce smaller sub-sequences or reverse complemented if necessary. \pb

The process of genome finishing may involve producing additional
oligonucleotide sequence to fill unknown regions in the scaffold. The scaffold
file format allows these additional insert sequences to be added to bridge
these gaps. These inserts can also be treated in a similar way to larger contig
sequences: trimmed and/or reverse complemented to match the corresponding gap
region. \pb

@@

The scaffold file is written using the YAML syntax. YAML is a data format
designed to be both machine readable and easy to edit manually. An example
scaffold file is shown in Figure 1. This file illustrates the text attributes
used to describe a scaffold and the joining of sequences in the genome build.
The basic layout of the scaffold file is a list - a series of entries where
each entry corresponds to a region of sequence in the generated scaffold
super-sequence. The right hand side indicates how the entries are joined
together. This figure is illustrates the putative scaffold and is not to scale.
\pb

@@

The first line of the scaffold file begins with three dashes. This is
a requirement to indicate the start of a YAML formatted document. The first
entry, highlighted in green, in the scaffold begins on the second line. The
first line of this entry begins with a dash character `-', dashes are used to
denote entries in a YAML list. The first line of each entry in the scaffold
therefore begins with a dash line. The next line is indented by two spaces
- whitespace is required to group similar attributes together. \pb

@@

The second entry in the scaffold, highlighted in orange, is identified by the
``unresolved'' tag, indicating a region of unknown sequence but known length.
The second line specifies the size of this unknown region. This entry therefore
inserts a region of 20 `N' characters in the scaffold. \pb

@@

This final entry also includes the ``inserts'' tag to add additional regions of
sequence. These inserts are also added as a YAML list with each insert
beginning with a dash line. The first insert, shown in purple, illustrates
using similar attributes to that of a sequence entry. The reverse, start and
stop tags, are used to trim and reverse complement the insert. Similarly the
`source' tag identifies the corresponding fasta sequence. The ``open'' and
``close'' tags are specific to inserts and determine where the insert is added
in the enclosing sequence. The region of the sequence inside these coordinates
is inclusively replaced by the specified insert fasta sequence. This is
highlighted in the putative scaffold in Figure 1 by the black lines bisecting
the blue sequence. \pb

The next insert, shown in brown, is specified using only the open position tag.
This illustrates that only one of either `open' or `close' tags is required.
When only one of these tag is used the corresponding opposing coordinate
position is calculated from the length of the insert fasta sequence. This
allows inserts to bridge into, and partially fill gap regions without requiring
an end coordinate position. \pb

@@

\paragraph{Reverse sort inserts:} If the sequence entry contains a list of
inserts, these inserts are reverse sorted by the insert close coordinate
position. The insert with the largest close position is first in the list and
the insert with the smallest close position is last. This preserves the
original intended coordinates as each insert is added to the sequence. \pb

@@

Scaffolder relies on both the Ruby programming language and the RubyGems
package management system. These are either already present or easily installed
on recent distributions of Mac OS X or Linux. The presence of both these tools
can be tested by typing the following commands on the terminal. \pb

@@

This code snippet illustrates that the scaffold object is initialised with two
arguments: the YAML loaded scaffold and the location of the fasta sequence
file. The scaffold is an enumerable array and the penultimate line illustrates
the trivial code required to generate the super-sequence. Further details of
the classes and methods related to each scaffold entry can be found in the
documentation for the Scaffolder::Region class in the documentation on the
Scaffolder website. \pb

@@

The aims of Scaffolder are twofold. The first is providing software which is
easy to install and simplifies the task of genome finishing. The second aim is
to facilitate reproducibility when joining large nucleotide sequences together
to create a genome sequence. \pb

@@

Scaffolder is designed to be as simple to use as possible. Assuming the Ruby
and RubyGems software are present, Scaffolder can be installed in a single
command line step. This should negate any possible barriers to entry which
require manual compilation of source code. \pb

Scaffolder uses a minimal and compact syntax to describe the genome scaffold.
This syntax is simple to construct and edit whilst being succinct and readable.
The goal of this syntax is to make genome scaffolds easy to construct in
a common text editor. Furthermore Scaffolder provides extensive documentation
both for the CLI and for the Ruby API. Therefore should any part of the
scaffolding process be unclear the documentation should make an solution easy
to find. \pb

@@

Configuring the final sequence in the scaffold file means the build is easier
to edit once already constructed. The YAML text format also allows comparison
of differences between scaffold builds using standard Unix tools such as diff.
This makes scaffold files easier to store in version control systems and allows
genome finishers methods similar to those in software development. \pb

@@

Scaffolder follows the Unix philosophy of ``Do one thing and do it well''. The
focus of Scaffolder is a simple to use tool for creating a genome scaffold. The
YAML data format for creating the scaffold file is standardised and easily
manipulated programmatically. This thereby means the scaffolding process
follows a second Unix tenet: ``If your data structures are good enough, the
algorithm to manipulate them should be trivial.''

@@

  \begin{description}
    \item[API:] Application programming interface
    \item[CLI:] Command line interface
    \item[PCR:] Polymerase chain reaction
    \item[YAML:] YAML ain't markup language\cite{yaml}
  \end{description}

@@

An example scaffold file written using the YAML syntax \cite{yaml} (left) and
the resulting putative scaffold (right). The scaffold contains three entries
and two inserts. Each entry in the scaffold file text is delimited by a `-' on
a new line and highlighted using separate colours. The scaffold diagram on the
right is not to scale but illustrates how the scaffold sequences are joined.
\pb
